---
title: "Final Individual Project"
author: "Daria Yakovleva"
date: "2023-11-15"
output: word_document
always_allow_html: true
---


```{r}
#Load libraries
library(readr)
library(ggplot2)
library(tidyverse)
library(rpart)
library(plyr)
library(dplyr)
library(rpart.plot)
library(maptree)
library(faraway)
library(gains)
library(liftr)
library(earth)
library(lattice)
library(Rprofet)
library(ROCR)
```


```{r}
#Load data
data <- read.csv("/Users/Dasha/Desktop/Individual final project/rent_data_training.csv")
#View(data)
```


```{r}
#Remove outliers
outlier_rent<- c(-6500, -13000,3500000, 4000000000000)
data <- data[!(data$Rent %in% outlier_rent), ]

#Calculate Z-scores for each data point
z_scores <- scale(data$Rent)

#Set a threshold for detecting outliers (adjust this as needed)
#A common threshold is |Z-score| > 3, which corresponds to about 99.7% of data falling within the threshold.
threshold <- 3

# Find outliers based on the Z-score
outliers <- which(abs(z_scores) > threshold, arr.ind = TRUE)

# Print the indices of the outliers
cat("Indices of Outliers:\n")
print(outliers)
```


```{r}
#Remove the outliers from the dataset

data_clean <- data[-outliers[,1], ]

#Print the cleaned dataset
#print(data_clean)
```


```{r}
# Assuming you have already cleaned the data and stored it in the 'data_clean' variable
# You can replace 'cleaned_data.csv' with your desired file name

# Save the cleaned data as a CSV file
write.csv(data_clean, file = "cleaned_data.csv", row.names = FALSE)

# Print a confirmation message
cat("Cleaned data saved as 'cleaned_data.csv'\n")

```


```{r}
data1<- read.csv("/Users/Dasha/Desktop/Individual final project/cleaned_data.csv")
#View(data1)
```


```{r}
#Finding rows with missing values
rows_with_missing_values <- which(!complete.cases(data1))

#Print the row indices with missing values
cat("Row indices with missing values:\n")
print(rows_with_missing_values)
```


#Barplot
```{r}
#Bar plot of Frequency of Properties by City
ggplot(data1, aes(x = City, fill = City)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Frequency of Properties by City", x = "City", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



```{r}
#Barplot of number of bathrooms
ggplot(data1, aes(x = as.factor(BHK), fill = BHK)) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-0.5) +
  labs(title = "Count of BHK", x = "BHK", y = "Count") +
  theme_minimal()
```


```{r}
#Box plot of rent and number of bathroom
data1_filtered <- data1[!(data1$Bathroom == 3 & data1$Rent > 100000), ]

ggplot(data1_filtered, aes(x = as.factor(Bathroom), y = Rent)) +
  geom_boxplot(fill = 'skyblue', color = 'black', alpha = 0.7) +
  labs(title = 'Rent Distribution by Number of Bathrooms', x = 'Number of Bathrooms', y = 'Rent') +
  theme_minimal()

```


#Histogram of the distribution of Rent
```{r}
ggplot(data1, aes(x = Rent)) +
  geom_histogram(binwidth = 8000, fill = "hotpink4", color = 'black') + 
  labs(x = "Rent", y = "Frequency", title = 'Distribution of the Rent')

# Logarithmic transformation of the 'Rent' column
log_Rent <- log(data1$Rent)
ggplot(data1, aes(x = log_Rent)) +
  geom_histogram(binwidth = 0.1, fill = "hotpink4", color = "black") + 
  labs(x = "Log Rent", y = "Frequency", title = "Histogram of Log-Transformed Rent")

#Scored Data
data1 <- cbind(data1, log_Rent)

```


#Boxplots
```{r}
#Boxplots of log_rent
par(mfrow = c(1, 1))
boxplot(log_Rent~City,data=data1, 
        main="Log of Rent by City", notch=TRUE,
        col=(c("hotpink3", "pink2", "skyblue2", "yellow2", "green2", "darkorange2")),
        xlab="City", ylab="Rent")

boxplot(log_Rent~Furnishing.Status,data=data1, main="Log of Rent by Furnishing Status", notch=TRUE,
        col=(c("hotpink3","pink2", "skyblue2")),
        xlab="Furnishing Status", ylab="Rent")

boxplot(log_Rent~BHK,data=data1, 
        main="Log of Rent by BHK", 
        col=(c("hotpink3", "pink2", "skyblue2", "yellow2", "green2", "darkorange2")),
        xlab="BHK", ylab="Rent")

boxplot(log_Rent~Area.Type,data=data1, 
        main="Log of Rent by Area Type",
        col=(c("hotpink3","pink2", "skyblue2")),
        xlab="Area Type", ylab="Rent")

boxplot(log_Rent~Tenant.Preferred,data=data1, 
        main="Log of Rent by Tenant Preferred", notch=TRUE,
        col=(c("hotpink3","pink2", "skyblue2")),
        xlab="Tenant Preferred", ylab="Rent")

boxplot(log_Rent~Point.of.Contact,data=data1, 
        main="Log of Rent by Point of Contact", notch=TRUE,
        col=(c("hotpink3","pink2", "skyblue2")),
        xlab="Point of Contact", ylab="Rent")

```


```{r}
#Boxplot of log_Rent by BHK and Furnishing Status
ggplot(data1, aes(x = as.factor(BHK), y = log_Rent, fill = Furnishing.Status)) +
  geom_boxplot() +
  labs(title = 'Boxplot of log_Rent by BHK and Furnishing Status', x = 'BHK', y = 'log_Rent') +
  theme_minimal()
```


#Count
```{r}
ggplot(data=data1) +
  geom_count(mapping = aes( x= BHK, y = Furnishing.Status, color = log_Rent))+
  scale_fill_gradient(low = "white", high = "navyblue")
```


#Scatter plot of log_Rent vs. Size. 
```{r}
#Scatter plot of log_Rent vs. size
ggplot(data1, aes(x = Size, y = log_Rent)) +
  geom_point(color = "navyblue", alpha = 0.7) +
  labs(title = 'Scatter Plot of Size and Rent', x = 'Size', y = 'Rent') +
  theme_minimal()
```



```{r}
# Plotting Bar Plots of Rent by Area Type and BHK
ggplot(data1, aes(x = Area.Type, y = log_Rent, fill = as.factor(BHK))) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  facet_wrap(~as.factor(BHK), scales = "free_y", ncol = 3) +
  labs(title = 'Bar Plot of Rent by Area Type and BHK', x = 'Area Type', y = 'Rent') +
  theme_minimal() +
  theme(strip.text.x = element_text(size = 10), strip.background = element_blank(), 
        axis.text.x = element_text(angle = 45, hjust = 1))  

```


#Correlation plot
```{r}
#Correlation plot
library(corrplot)
cor_matrix<-cor(data1[,c(2, 3, 10, 13)])

par(mar = c(2, 2, 2, 2))
corrplot(cor_matrix, addCoef.col = TRUE,number.cex=0.75)
```



```{r}
#Remove entries with rent above 300,000 and when BHK = 1 and Bathroom = 1
data1$Area.Type[data1$Area.Type == "Built Area"] <- "Carpet Area"
filtered_data <- subset(data1, Rent <= 300000 & BHK == 1 & Bathroom == 1)

#Use aggregate to calculate the average rent for each city in the filtered data
average_rent_by_city <- aggregate(Rent ~ City, data = filtered_data, FUN = function(x) round(mean(x, na.rm = TRUE)))

#Display the result
print(average_rent_by_city)
```


#Association rules
```{r include=FALSE}
#Load libraries
library(arules)
library(arulesViz)
```


```{r}
#Making a new variable "cheap_rent" based on Rent column and choosing breaks
x <- 5700
breaks2 <- c(-13000, x , 100000)

#Bin data 
filtered_data$cheap_rent <- cut(filtered_data$Rent, breaks = breaks2, labels = c("Low", "High")) 
```


```{r}
#Choosing only categorical variables for the association rules
cat_var <- filtered_data[, c("Floor", "Area.Type", "Area.Locality", "City", "Furnishing.Status", "Tenant.Preferred", "Point.of.Contact", "cheap_rent")]
str(cat_var)
```


Also, I converted categorical columns to factors. 
```{r include=FALSE}
#Convert all categorical columns to factors
cat_var <- as.data.frame(lapply(cat_var, factor))
cat_var <- transactions(cat_var)
itemLabels(cat_var)

#Check the structure of the updated data frame
str(cat_var)
```


```{r}
#Perform association rules
rules <- apriori(cat_var)
inspect(rules)
```


```{r}
itemFrequencyPlot(cat_var, topN=10, type='absolute',horiz=T,cex=0.7)
```


```{r include = FALSE}
#Set rhs=c("cheap_rent=Low")
rules <- apriori(cat_var, parameter = list(minlen=3, supp = .05, conf = .05),
                 appearance = list(rhs=c("cheap_rent=Low"), default = "lhs"),
                 control = list(verbose=T))
inspect(rules)
```



```{r include = FALSE}
#sort rules
rules.sorted <- sort(rules, by="confidence")
inspect(rules.sorted)
```


Now, we find redundant rules and remove them. 
```{r}
#Find redundant rules
is.redundant(rules.sorted)

rules.1 <- rules.sorted[!is.redundant(rules.sorted)]
inspect(rules.1)
```



## Visualizing association rules
Plotting the association rules. 
```{r}
plot(rules.1)
plot(rules.1, method = "graph", control=list(type="items"))
```


Now, I am sorting the rules by confidence in descending order and chose the top 10 rules. Then, I plot the results for the rules for the three new variables that I have created in the beginning of the association rule. 

```{r}
#Fancy Plot
plot(rules.1, method = "graph", measure = "support", shading = "confidence",
     engine = 'visNetwork')
```


#Spliting data set into Train and Test datasets
```{r}
#Splitting the dataset into the Training set and Test set (80%/20%)
set.seed(123)
n = dim(data1)[1]
split = sample(1:n,size = n*0.8, replace = F)

#Train set
train_set = data1[split,] 

#Test set
test_set = data1[-split,]
```


#Draw the tree and predict
```{r}
#Fitting Decision Tree Regression to the dataset
tree = rpart(log_Rent ~ BHK + Size + Area.Type + City + Furnishing.Status + Tenant.Preferred + Bathroom + Point.of.Contact, 
             data = train_set,
             cp = 0.002,
             minbucket = 150)

#Prune Tree
pruned = clip.rpart(tree, best = 7)
```


```{r}
#Draw tree
draw.tree (pruned, cex=0.8, 
           nodeinfo = TRUE, units = "log_Rent",
           cases ="obs",
           digits = 3, print.levels = TRUE,
           new = TRUE)

#Predicting a new result with Decision Tree Regression
tree_pred = predict(pruned, newdata = test_set) 
```


#Regression Model
```{r}
#Regression Model
reg_model_1 <- lm(log_Rent ~ BHK + Size + Area.Type + City + Furnishing.Status + Tenant.Preferred + Bathroom + Point.of.Contact, data = train_set)
summary(reg_model_1)
```


#Regression Model
```{r}
#Remove area type since it is not very significant
reg_model_2 <- lm(log_Rent ~ BHK + Size + City + Furnishing.Status + Tenant.Preferred + Bathroom + Point.of.Contact, data = train_set)
summary(reg_model_2)

reg_pred<- predict(reg_model_2, newdata = test_set, type = "response")
```


#MARS Model
```{r}
train_set <- na.omit(train_set)
mars <- earth(log_Rent ~ BHK + Area.Type + Size + City + Furnishing.Status + Tenant.Preferred + Bathroom + Point.of.Contact, data = train_set)
summary(mars)

mars_pred <- predict(mars, newdata = test_set, type = "response")
```


#Gains tables for the three models
#Tree
```{r}
#Gains for tree
treegains = gains(test_set$log_Rent, tree_pred)


#Extract columns from gains object into data frame
treegains_data <- data.frame(depth = treegains$depth,
                           mean.prediction = treegains$mean.prediction,
                           mean.response = treegains$mean.resp)

#Plot of Depth against mean responses
ggplot(treegains_data, aes(x = depth)) +
  geom_line(aes(y = mean.prediction, color = "Predicted"), linetype = "dashed") +
  geom_line(aes(y = mean.response, color = "Actual")) +
  labs(x = "Depth", y = "Mean Response",
       title = "Mean Response by Sampled Population for Tree Model") +
  scale_color_manual(name = "Response", values = c("Predicted" = "blue", 
                                                   "Actual" = "red")) +
  theme_minimal()
```


```{r}
# Data frame with actual and predicted values
tree_results <- data.frame(log_Rent = test_set$log_Rent, 
                            Predicted = tree_pred)
```


#Regression
```{r}
# Gains for Regression Model
reggains = gains(test_set$log_Rent, reg_pred)
 

# Extract columns from gains object into data frame
reggains_data <- data.frame(depth = reggains$depth,
                           mean.prediction = reggains$mean.prediction,
                           mean.response = reggains$mean.resp)

# Plot of Depth against mean responses
ggplot(reggains_data, aes(x = depth)) +
  geom_line(aes(y = mean.prediction, color = "Predicted"), linetype = "dashed") +
  geom_line(aes(y = mean.response, color = "Actual")) +
  labs(x = "Depth", y = "Mean Response",
       title = "Mean Response by Sampled Population for Regression Model") +
  scale_color_manual(name = "Response", values = c("Predicted" = "blue", 
                                                   "Actual" = "red")) +
  theme_minimal()
```

#MARS Model
```{r} 
# Gains for MARS Model
marsgains = gains(test_set$log_Rent, mars_pred)
 

# Extract columns from gains object into data frame
marsgains_data <- data.frame(depth = marsgains$depth,
                           mean.prediction = marsgains$mean.prediction,
                           mean.response = marsgains$mean.resp)

# Plot of Depth against mean responses
ggplot(marsgains_data, aes(x = depth)) +
  geom_line(aes(y = mean.prediction, color = "Predicted"), linetype = "dashed") +
  geom_line(aes(y = mean.response, color = "Actual")) +
  labs(x = "Depth", y = "Mean Response",
       title = "Mean Response by Sampled Population for MARS Model") +
  scale_color_manual(name = "Response", values = c("Predicted" = "blue", 
                                                   "Actual" = "red")) +
  theme_minimal()
```
# Prediction
#Tree Prediction
```{r}
# Data frame with actual and predicted values
tree_results <- data.frame(log_Rent = test_set$log_Rent, 
                            Predicted = tree_pred)
```


```{r}
#Regression
#Data frame with actual and predicted values
reg_results <- data.frame(log_Rent = test_set$log_Rent, 
                            Predicted = reg_pred)

reg_eval <- (1 / nrow(reg_results)) * sum(abs(exp(reg_results$log_Rent) - exp(reg_results$Predicted)) / exp(reg_results$log_Rent))
print(reg_eval)
```


```{r}
#Tree
#Calculate the metric for the training data
tree_eval <- (1 / nrow(tree_results)) * sum(abs(exp(tree_results$log_Rent) - exp(tree_results$Predicted)) / exp(tree_results$log_Rent))
print(tree_eval)

```


```{r}
#MARS
#Data frame with actual and predicted values
mars_results <- data.frame(log_Rent = test_set$log_Rent, 
                            Predicted = mars_pred)

mars_eval <- (1 / nrow(mars_results)) * sum(abs(exp(mars_results$log_Rent) - exp(mars_pred)) / exp(mars_results$log_Rent))
print(mars_eval)
```


```{r}
test_data<- read.csv("/Users/Dasha/Desktop/Individual final project/rent_data_test.csv")
```


```{r}
mars_pred2 <- predict (mars, newdata = test_data, type = "response")

Predict_Rent <- exp(mars_pred2)

Predict_Rent_data <- cbind(test_data, Predict_Rent)
write.csv(Predict_Rent_data, "Predicted_Rent.csv")
```

